\section{Conclusions} \label{sec:conclusion}

Applying machine learning to compile-time and runtime optimizations requires generating features first. This is a time consuming process, it needs supervision by an expert, and even then we cannot be sure that the selected features are optimal. In this paper we present a novel tool for building optimization heuristics, \DeepTune, which forgoes feature extraction entirely, relying on powerful language modeling techniques to automatically build complex and effective representations of programs directly from raw source code. The result translates into a huge reduction in development effort, improved heuristic performance, and more simple model designs.

Our approach is fully automated. Using \DeepTune, compiler developers no longer need to spend months using statistical methods and profile counters to select program features via trial and error.  It is worth mentioning that we do not tailor our model design or parameters for the optimization task at hand, yet we achieve performance on par with and in most cases \emph{exceeding} state-of-the-art predictive models.

We used \DeepTune to automatically construct heuristics for two challenging optimization problems: selecting the optimal execution device for OpenCL kernels, and selecting OpenCL thread coarsening factors. In both cases, we outperform state-of-the-art predictive models, achieving performance improvements of 16\% and 12\%, respectively. We have also shown that the \DeepTune architecture allows us to exploit information learned from another optimization problem to give the learning a boost. Doing so provides up to a 16\% performance improvement when training using a handful of training programs. We suspect that this approach will be useful for other optimization tasks for which training programs are a scarce resource.

In future work, we will extend our heuristic construction approach by automatically learning dynamic features over raw data; apply unsupervised learning techniques~\cite{Le2012} over unlabeled source code to further improve learned representations of programs; \review{and deploy trained \DeepTune heuristic models to low power embedded systems using optimization and compression of neural networks~\cite{Han2015}}.
