\section{Introduction}

% Paper abstract.
%
%We introduce DeepSmith, a novel machine learning approach to accelerating
%compiler validation through the inference of generative models for compiler
%inputs. Our approach \emph{infers} a learned model of the structure of real
%world code based on a large corpus of open source code. Then, it uses the
%model to automatically generate tens of thousands of realistic programs.
%Finally, we apply established differential testing methodologies on them to
%expose bugs in compilers. We apply our approach to the OpenCL programming
%language, automatically exposing bugs with little effort on our side. In 1,000
%hours of automated testing of commercial and open source compilers, we
%discover bugs in all of them, submitting 67 bug reports. Our test cases are on
%average two orders of magnitude smaller than the state-of-the-art, require
%$3.03\times$ less time to generate and evaluate, and expose bugs which the
%state-of-the-art cannot. Our random program generator, comprising only 500
%lines of code, took 12 hours to train for OpenCL versus the state-of-the-art
%taking 9 man months to port from a generator for C and 50,000 lines of code.
%With 18 lines of code we extended our program generator to a second language,
%uncovering crashes in Solidity compilers in 12 hours of automated testing.

Compilers should produce correct code for valid inputs, and meaningful errors for invalid inputs. Failure to do so can hinder software development or even cause catastrophic runtime errors. Still, properly testing compilers is hard. Modern optimising compilers are large and complex programs, and their input space is huge. Hand designed suites of test programs, while important, are inadequate for covering such a large space and will not touch all parts of the compiler.

Random test case generation --- \emph{fuzzing} --- is a well established and effective method for identifying compiler bugs~\cite{Chen2014a,Chen2013,Kossatchev2005}. When fuzzing, randomly generated valid or semi-valid inputs are fed to the compiler. Any kind of unexpected behaviour, including crashes, freezes, or wrong binaries, indicates a compiler bug. While crashes and freezes in the compiler are easy to detect, determining that binaries are correctly compiled is not generally possible without either developer provided validation for the particular program's behaviour or a gold standard compiler from which to create reference outputs. In the absence of those, Differential Testing~\cite{McKeeman1998} can be used. The generated code and a set of inputs form a \emph{test case} which is compiled and executed on multiple \emph{testbeds}. If the test case should have deterministic behaviour, but the output differs between testbeds, then a bug has been discovered.

Compiler fuzzing requires efficiently generating test cases that trigger compiler bugs. The state-of-the-art approach, CSmith~\cite{Yang2011}, generates large random programs by defining and sampling a probabilistic grammar which covers a subset of the C programming language. Through this grammar, CSmith ensures that the generated code easily passes the compiler front-end and stresses the most complex part of the compiler, the middle-end. Complex static and dynamic analyses make sure that programs are free from undefined behaviour. The programs are then differentially tested.

While CSmith has been successfully used to identify hundreds of bugs in otherwise-robust  compilers, it and similar approaches have a significant drawback. They represent a huge undertaking and require a thorough understanding of the target programming language. CSmith was developed over the course of years, and consists of over 41k lines of handwritten C++ code. By tightly coupling the generation logic with the target programming language, each feature of the grammar must be painstakingly and expertly engineered for each new target language. For example, lifting CSmith from C to OpenCL~\cite{Lidbury2015a} --- a superficially simple task --- took 9 months and an additional 8k lines of code. Given the difficulty of defining a new grammar, typically only a subset of the language is implemented.

This chapter introduces \emph{DeepSmith}, a novel machine learning approach to accelerating compiler validation through the inference of generative models for compiler inputs. DeepSmith is a fast, effective, and low effort approach to the generation of random programs for compiler fuzzing. The methodology uses recent advances in deep learning to automatically construct probabilistic models of how humans write code, instead of painstakingly defining a grammar to the same end. By training a deep neural network on a corpus of handwritten code, it is able to infer both the syntax and semantics of the programming language and the common constructs and patterns. The approach essentially frames the generation of random programs as a language modelling problem. This greatly simplifies and accelerates the process. The expressiveness of the generated programs is limited only by what is contained in the corpus, not the developer's expertise or available time. Such a corpus can readily be assembled from open source repositories.

This work primarily targets OpenCL, an open standard for programming heterogeneous systems, though the approach is largely language agnostic. OpenCL is chosen for three reasons: it is an emerging standard with the challenging promise of functional portability across a diverse range of heterogeneous hardware; OpenCL is compiled ``online'', meaning that even compiler crashes and freezes may not be discovered until a product is deployed to customers; and there is already a hand written random program generator for the language to compare against.

This chapter is organised as follows:  Section~\ref{sec:deepsmith} presents DeepSmith, a novel approach to compiler validation. Section~\ref{sec:deepsmith-experimental-setup} describes the experimental setup of an extensive evaluation of OpenCL compilers using DeepSmith. Section~\ref{sec:deepsmith-eval} evaluates the results of the experiment, with Section~\ref{subsec:deepsmith-solidity-extensibility} containing preliminary results supporting DeepSmith's potential for multi-lingual compiler fuzzing. Section~\ref{sec:deepsmith-conclusion} provides concluding remarks for this chapter.