\section{Program Optimisation}
\label{sec:related-work-optimisation}

Modern compilers are complex, frequently containing hundreds of optimisation passes. Determining which optimisation passes should be applied, and in what order, is a challenge that depends on a variety of factors from the properties of the program being compiled to the target hardware. Current state-of-practise is for compilers to use a fixed ordering of optimisations, and for each optimisation to contain a heuristic to determine when with what parameters it should be used, and under what circumstances. Such heuristics are typically developed by hand at the expense of great effort and compiler expertise.

Unfortunately, the complex interactions between hardware and software often cannot be captured by heuristic and analytical methods, resulting in performance being left on the table.

Iterative compilation is a method of performance tuning in which a program is compiled and profiled using multiple different configurations of optimisations in order to find the configuration which maximises performance. Unlike analytical methods which attempt to predict the parameters that produce good performance, iterative compilation is empirical. A set of candidate configurations are selected, and for each, the program is compiled and profiled. The configuration that maximises some metric for reward (such as speedup) is selected. In one of the first works of the field, \citeauthor{Bodin1998} demonstrated the utility of the approach at finding good configurations in the non-linear three-parameter optimisation space of a matrix multiplication benchmark~\cite{Bodin1998}. Iterative compilation has since been demonstrated to be a highly effective form of empirical performance tuning for selecting compiler optimisations.

While conceptually simple, the empirical nature of iterative compilation yields good results.

PetaBricks is a language and compiler for algorithmic choice~\cite{Ansel2009a}. Users provide multiple implementations of algorithms, optimised for different parameters or use cases. This creates a search space of possible execution paths for a given program. This has been combined with autotuning techniques for enabling optimised multigrid programs~\cite{Chan2009}, with the wider ambition that these autotuning techniques may be applied to all algorithmic choice programs~\cite{Ansel2014}. While this helps produce efficient programs, it places a great burden on the developer, requiring them to provide enough contrasting implementations to make a search of the optimisation space fruitful.

\todo{more algorithmic choice~\cite{Hutter2009a}}

\todo{domain-specific algorithmic choice~\cite{Whaley1998}}

One of the challenges of iterative compilation is the exponential blow up of optimisation space size when considering many independent parameters. Many compilers contain dozens or hundreds of discrete optimisation transformations, rendering an exhaustive search of the optimisation space infeasible. This has driven the development of methods for reducing the cost of evaluating configurations. These methods reduce evaluation costs either by shrinking the dimensionality or size of the optimisation space, or by guiding a directed search to traverse a subset of the space.

\todo{machine learning methods to reduce the iterative compilation search space are discussed in Section~\ref{sec:related-work-machine-learning-for-pl}.}

In~\cite{Saclay2010,Memon2013,Fursin2014}, \citeauthor{Fursin2014} advocate a ``big data'' driven approach to autotuning, arguing that the challenges facing widespread adoption of autotuning and machine learning methodologies can be attributed to: a lack of common, diverse benchmarks and data sets; a lack of common experimental methodology; problems with continuously changing hardware and software stacks; and the difficulty to validate techniques due to a lack of sharing in publications. They propose a system for addressing these concerns, the Collective Mind knowledge system, which provides a modular infrastructure for sharing autotuning performance data and related artefacts across the internet.

% Georgiou, K., Blackmore, C., Xavier-de-Souza, S., & Eder, K. (2018). Less is More: Exploiting the Standard Compiler Optimization Levels for Better Performance and Energy Consumption. In SCOPES.
\todo[inline]{Using \emph{less} flags that -O2 can improve performance~\cite{Georgiou2018}.}

Machine learning has been successful applied to this problem,
in~\cite{Stephenson2003}, using ``meta optimisation'' to tune compiler
heuristics through an evolutionary algorithm to automate the search of
the optimisation space. \citeauthor{Fursin2011} continued this with
Milepost GCC, the first machine learning-enabled self-tuning
compiler~\cite{Fursin2011}. A recent survey of the use of machine
learning to improve heuristics quality by \citeauthor{Burke2013}
concludes that the automatic \emph{generation} of these self-tuning
heuristics but is an ongoing research challenge that offers the
greatest generalisation benefits~\cite{Burke2013}.

% Ashouri, A. H., Killian, W., Cavazos, J., Palermo, G., & Silvano, C. (2018). A Survey on Compiler Autotuning using Machine Learning. CSUR, 51(5). https://doi.org/10.1145/3197978
\todo[inline]{Autotuning through ML survey~\cite{Ashouri2018}.}

An approach to online tuning of parallel programs is presented in~\cite{Ansel2012} which partitions the available parallel resources of a device in to two partitions and then executes two different configurations simultaneously using each partition. The configuration used for one of the configuration is guaranteed to be ``safe'', and the performance

 % Eastep, J., Wingate, D., & Agarwal, A. (2011). Smart Data
 % Structures: An Online Machine Learning Approach to Multicore Data
 % Structures. In Proceedings of the 8th ACM International Conference
 % on Autonomic Computing (pp. 11–20). New York, NY, USA:
 % ACM. doi:10.1145/1998582.1998587
Online reinforcement learning for optimising data structures online, \cite{Tesauro2005}.

 % Tesauro, G. (2005). Online Resource Allocation Using Decompositional
 % Reinforcement Learning. In AAAI (pp. 886–891).
Reinforcement learning for resource allocation~\cite{Eastep2011}.

% W. F. Ogilvie, P. Petoumenos, Z. Wang, and H. Leather, “Intelligent
% Heuristic Construction with Active Learning,” in 18th International
% Workshop on Compilers for Parallel Computing, 2015.
\todo[inline]{Using Active Learning to speed up the learning of compiler heuristics~\cite{Ogilvie2015}. Towards online autotuning, albeit only with binary optimisation parameter.}

% Ashouri, A. H., Milano, G., Palermo, G., Park, E., Cavazos, J., & Silvano, C. (2016). COBAYN: Compiler Autotuning Framework Using Bayesian Networks. TACO, 13(2).
\todo[inline]{Amir's ML autotuning framework~\cite{Ashouri2016}.}

%
% SOME EXAMPLES OF ML IN THE WILD:
%

In~\cite{Saclay2010,Memon2013,Fursin2014}, \citeauthor{Fursin2014} advocate a collaborative and ``big data'' driven approach to autotuning, arguing that the challenges facing the widespread adoption of autotuning and machine learning methodologies can be attributed to: a lack of common, diverse benchmarks and data sets; a lack of common experimental methodology; problems with continuously changing hardware and software stacks; and the difficulty to validate techniques due to a lack of sharing in publications. They propose a system for addressing these concerns, the Collective Mind knowledge system, which, while in early stages of ongoing development, is intended to provide a modular infrastructure for sharing autotuning performance data and related artefacts across the internet. In addition to sharing performance data, the approach taken in this thesis emphasises the collective \emph{exploitation} of such performance data, so that data gathered from one device may be used to inform the autotuning decisions of another. This requires each device to maintain local caches of shared data to remove the network overhead that would be present from querying a single centralised data store during execution of a hot path. The current implementation of Collective Mind uses a NoSQL JSON format for storing performance data. The relational schema used in this thesis offers greater scaling performance and lower storage overhead as the amount of performance data grows.

% Kraska, T., Beutel, A., Chi, E. H., Dean, J., & Polyzotis, N. (2017). The Case for Learned Index Structures. ArXiv:1712.01208.
\todo[inline]{The Case for Learned Index Structures~\cite{Kraska2017}.}

Whereas iterative compilation requires an expensive offline training phase to search an optimisation space, dynamic optimisers perform this optimisation space exploration at runtime, allowing programs to respond to dynamic features ``online''. This is a challenging task, as a random search of an optimisation space may result in configurations with vastly sub-optimal performance. In a real world system, evaluating many sub-optimal configurations will cause a significant slowdown of the program. Thus a requirement of dynamic optimisers is that convergence time towards optimal parameters is minimised.

Existing dynamic optimisation research has typically taken a low level approach to performing optimisations. Dynamo is a dynamic optimiser which performs binary level transformations of programs using information gathered from runtime profiling and tracing~\cite{Bala2000}. While this provides the ability to respond to dynamic features, it restricts the range of optimisations that can be applied to binary transformations. These low level transformations cannot match the performance gains that higher level parameter tuning produces.

An interesting related tangent to iterative compilation is the development of so-called ``super-optimisers''. In~\cite{Massalin1987}, the smallest possible program which performs a specific function is found through a brute force enumeration of the entire instruction set. Starting with a program of a single instruction, the super-optimiser tests to see if any possible instruction passes a set of conformity tests. If not, the program length is increased by a single instruction and the process repeats. The exponential growth in the size of the search space is far too expensive for all but the smallest of hot paths, typically less than 13 instructions. The optimiser is limited to register to register memory transfers, with no support for pointers, a limitation which is addressed in~\cite{Joshi2002}. Denali is a super-optimiser which uses constraint satisfaction and rewrite rules to generate programs which are \emph{provably} optimal, instead of searching for the optimal configuration through empirical testing. Denali first translates a low level machine code into guarded multi-assignment form, then uses a matching algorithm to build a graph of all of a set of logical axioms which match parts of the graph, before using boolean satisfiability to disprove the conjecture that a program cannot be written in $n$ instructions. If the conjecture cannot be disproved, the size of $n$ is increased and the process repeats.


% \section{Performance Tuning for Heterogeneous Parallelism}

The complex interactions between optimisations and heterogeneous hardware makes performance tuning for heterogeneous parallelism a difficult task. Performant GPGPU programs require careful attention from the developer to properly manage data layout in DRAM, caching, diverging control flow, and thread communication. The performance of programs depends heavily on fully utilising zero-overhead thread scheduling, memory bandwidth, and thread grouping. \citeauthor{Ryoo2008a} illustrate the importance of these factors by demonstrating speedups of up to $432\times$ for matrix multiplication in CUDA by appropriate use of tiling and loop unrolling~\cite{Ryoo2008a}. The importance of proper exploitation of local shared memory and synchronisation costs is explored in~\cite{Lee2010}.

In~\cite{Chen2014}, data locality optimisations are automated using a description of the hardware and a memory-placement-agnostic compiler. The authors demonstrate impressive speedups of up to $2.08\times$, although at the cost of requiring accurate memory hierarchy descriptor files for all targeted hardware. The descriptor files must be hand generated, requiring expert knowledge of the underlying hardware in order to properly exploit memory locality.

Data locality for nested parallel patterns is explored in~\cite{Lee}. The authors use an automatic mapping strategy for nested parallel skeletons on GPUs, which uses a custom intermediate representation and a CUDA code generator, achieving $1.24\times$ speedup over hand optimised code on 7 of 8 Rodinia benchmarks.

Reduction of the GPGPU optimisation space is demonstrated in~\cite{Ryoo2008}, using the common subset of optimal configurations across a set of training examples. This technique reduces the search space by 98\%, although it does not guarantee that for a new program, the reduced search space will include the optimal configuration.

\citeauthor{Magni2014} demonstrated that thread coarsening of OpenCL kernels can lead to speedups in program performance between $1.11\times$ and $1.33\times$ in~\cite{Magni2014}. The authors achieve this using a machine learning model to predict optimal thread coarsening factors based on the static features of kernels, and an LLVM function pass to perform the required code transformations.

A framework for the automatic generation of OpenCL kernels from high-level programming concepts is described in~\cite{Steuwer2015}. A set of rewrite rules is used to transform high-level expressions to OpenCL code, creating a space of possible implementations. This approach is ideologically similar to that of PetaBricks, in that optimisations are made through algorithmic choice, although in this case the transformations are performed automatically at the compiler level. The authors report performance on a par with that of hand written OpenCL kernels.


% \section{Autotuning Algorithmic Skeletons}

An enumeration of the optimisation space of Intel Thread Building Blocks in~\cite{Contreras2008} shows that runtime knowledge of the available parallel hardware can have a significant impact on program performance. \citeauthor{Collins2012} exploit this in~\cite{Collins2012}, first using Principle Components Analysis to reduce the dimensionality of the space of possible optimisation parameters, followed by a search of parameter values to optimise program performance by a factor of $1.6\times$ over values chosen by a human expert. In~\cite{Collins2013}, they extend this using static feature extraction and nearest neighbour classification to further prune the search space, achieving an average 89\% of the oracle performance after evaluating 45 parameters.

\citeauthor{Dastgeer2011} developed a machine learning based auto-tuner for the SkePU skeleton library in~\cite{Dastgeer2011}. Training data is used to predict the optimal execution device (i.e.\ CPU, GPU) for a given program by predicting execution time and memory copy overhead based on problem size. The auto-tuner only supports vector operations, and there is limited cross-architecture evaluation. In~\cite{Dastgeer2015a}, the authors extend SkePU to improve the data consistency and transfer overhead of container types, reporting up to a $33.4\times$ speedup over the previous implementation.


% \section{Code Generation and Autotuning for Stencils}

Stencil codes have a variety of computationally expensive uses from fluid dynamics to quantum mechanics. Efficient, tuned stencil kernels are highly sought after, with early work in \citeyear{Bolz2003} by \citeauthor{Bolz2003} demonstrating the capability of GPUs for massively parallel stencil operations~\cite{Bolz2003}. In the resulting years, stencil codes have received much attention from the performance tuning research community.

\citeauthor{Ganapathi2009} demonstrated early attempts at autotuning multi-core stencil codes in~\cite{Ganapathi2009}, drawing upon the successes of statistical machine learning techniques in the compiler community, as discussed in Section~\ref{sec:iterative-compilation}. They present an auto-tuner which can achieve performance up to 18\% better than that of a human expert. From a space of 10 million configurations, they evaluate the performance of a randomly selected 1500 combinations, using Kernel Canonical Correlation Analysis to build correlations between tunable parameter values and measured performance targets. Performance targets are L1 cache misses, TLB misses, cycles per thread, and power consumption. The use of KCAA restricts the scalability of their system as the complexity of model building grows exponentially with the number of features. In their evaluation, the system requires two hours of compute time to build the KCAA model for only 400 seconds of benchmark data. They present a compelling argument for the use of energy efficiency as an optimisation target in addition to runtime, citing that it was the power wall that lead to the multi-core revolution in the first place. Their choice of only 2 benchmarks and 2 platforms makes the evaluation of their auto-tuner somewhat limited.

\citeauthor{Berkeley2009} targeted 3D stencils code performance in~\cite{Berkeley2009}. Stencils are decomposed into core blocks, sufficiently small to avoid last level cache capacity misses. These are then further decomposed to thread blocks, designed to exploit common locality threads may have within a shared cache or local memory. Thread blocks are divided into register blocks in order to take advantage of data level parallelism provided by the available registers. Data allocation is optimised on NUMA systems. The performance evaluation considers speedups of various optimisations with and without consideration for host/device transfer overhead.

\citeauthor{Kamil2010} present an autotuning framework in~\cite{Kamil2010} which accepts as input a Fortran 95 stencil expression and generates tuned shared-memory parallel implementations in Fortan, C, or CUDA. The system uses an IR to explore autotuning transformations, enumerating a subset of the optimisation space and recording only a single execution time for each configuration, reporting the fastest. They demonstrate their system on 4 architectures using 3 benchmarks, with speedups of up to $22\times$ compared to serial implementations. The CUDA code generator does not optimise for the GPU memory hierarchy, using only global memory. As demonstrated in this thesis, improper utilisation of local memory can hinder program performance by two orders of magnitude. There is no directed search or cross-program learning.

In~\cite{Zhang2013a}, \citeauthor{Zhang2013a} present a code generator
and auto-tuner for 3D Jacobi stencil codes. Using a DSL to express
kernel functions, the code generator performs substitution from one of
two CUDA templates to create programs for execution on GPUs. GPU
programs are parameterised and tuned for block size, block dimensions,
and whether input data is stored in read only texture memory. This
creates an optimisation space of up to 200 configurations. In an
evaluation of 4 benchmarks, the authors report impressive performance
that is comparable with previous implementations of iterative Jacobi
stencils on GPUs~\cite{Holewinski2012, Phillips2010}. The dominating
parameter is shown to be block dimensions, followed by block size,
then read only memory. The DSL presented in the paper is limited to
expressing only Jacobi Stencils applications. Critically, their
auto-tuner requires a full enumeration of the parameter space for each
program. Since there is no indication of the compute time required to
gather this data, it gives the impression that the system would be
impractical for the needs of general purpose stencil computing. The
auto-tuner presented in this thesis overcomes this drawback by learning
parameter values which transfer to unseen stencils, without the need
for an expensive tuning phase for each program and architecture.

In~\cite{Christen2011}, \citeauthor{Christen2011} present a DSL for
expressing stencil codes, a C code generator, and an auto-tuner for
exploring the optimisation space of blocking and vectorisation
strategies. The DSL supports stencil operations on arbitrarily
high-dimensional grids. The auto-tuner performs either an exhaustive,
multi-run Powell search, Nelder Mead, or evolutionary search to find
optimal parameter values. They evaluate their system on two CPUS and
one GPU using 6 benchmarks. A comparison of tuning results between
different GPU architectures would have been welcome, as the results
presented in this thesis show that devices have different responses to
optimisation parameters. The authors do not present a ratio of the
available performance that their system achieves, or how the
performance of optimisations vary across benchmarks or devices.

A stencil grid can be decomposed into smaller subsections so that
multiple GPUs can operate on each subsection independently. This
requires a small overlapping region where each subsection meets ---
the halo region --- to be shared between devices. For iterative
stencils, values in the halo region must be synchronised between
devices after each iteration, leading to costly communication
overheads. One possible optimisation is to deliberately increase the
size of the halo region, allowing each device to compute updated
values for the halo region, instead of requiring a synchronisation of
shared state. This reduces the communication costs between GPUs, at
the expense of introducing redundant computation. Tuning the size of
this halo region is the goal of PARTANS~\cite{Lutz2013}, an autotuning
framework for multi-GPU stencil computations. \citeauthor{Lutz2013}
explore the effect of varying the size of the halo regions using six
benchmark applications, finding that the optimal halo size depends on
the size of the grid, the number of partitions, and the connection
mechanism (i.e.\ PCI express). The authors present an auto-tuner which
determines problem decomposition and swapping strategy offline, and
performs an online search for the optimal halo size. The selection of
overlapping halo region size compliments the selection of work-group
size which is the subject of this thesis. However, PARTANS uses a
custom DSL rather than the generic interface provided by SkelCL, and
PARTANS does not learn the results of tuning across programs, or
across multiple runs of the same program.

% \section{Compilers}

\todo[inline]{Using active learning to minimise the cost of iterative compilation~\cite{Ogilvie2017}.}

\todo[inline]{Continuous learning of compiler heuristics~\cite{Tartara2013,Fursin2011}.}

Section~\ref{subsec:training-with-synthetic-benchmarks} describes training with synthetic benchmarks.

\subsection{Program Representations: Features and Embeddings}

\todo[inline]{Automatic feature \emph{selection} has been done before. The technique proposed in this thesis is for feature \emph{generation}, i.e. extracting a novel representation from text.}

Machine learning has emerged as a viable means in automatically constructing heuristics for code optimisation~\cite{Stephenson2005,Agakov,Wang2010,Kulkarni2012,Ding2015,Muralidharan2016}. Its great advantage is that it can adapt to changing hardware platforms as it has no a priori assumptions about their behaviour. The success of machine learning based code optimisation has required having a set of high-quality features that can capture the important characteristics of the target program. Given that there is an infinite number of these potential features, finding the right set of features is a non-trivial, time-consuming task.

Various forms of program features have been used in compiler-based machine learning. These include static code structures~\cite{Jiang2010} and runtime information such as system load~\cite{Wen2015} and performance counters~\cite{Dubach2009}. In compiler research, the feature sets used for predictive models are often provided without explanation and rarely is the quality of those features evaluated. More commonly, an initial large, high dimensional candidate feature space is pruned via feature selection~\cite{Stephenson2005}, or projected into a lower dimensional space~\cite{Collins2013,Dubach2007}. FEAST employs a range of existing feature selection methods to select useful candidate features~\cite{Ting2016}. Unlike these approaches, DeepTune extracts features and reduces the dimensionality of the feature space completely internally and without expert guidance.

Park \emph{et al.} present a unique graph-based approach for feature representations~\cite{Park2012}. They use a Support Vector Machine where the kernel is based on graph similarity metric. Their technique still requires hand coded features at the basic block level, but thereafter, graph similarity against each of the training programs takes the place of global features. Being a kernel method, it requires that training data graphs be shipped with the compiler, which may not scale as the size of the training data grows with the number of instances, and some training programs may be very large. Finally, their graph matching metric is expensive, requiring $O(n^3)$ to compare against each training example. By contrast, our method does not need any hand built static code features, and the deployment memory footprint is constant and prediction time is linear in the length of the program, regardless of the size of the training set.

A few methods have been proposed to automatically generate features from the compiler's intermediate representation~\cite{Namolaru2010a,Leather2014}. These approaches closely tie the implementation of the predictive model to the compiler IR, which means changes to the IR will require modifications to the model. The work of \cite{Leather2014} uses genetic programming to search for features, and required a huge grammar to be written, some 160kB in length. Although much of this can be created from templates, selecting the right range of capabilities and search space bias is non trivial and up to the expert. The work of \cite{Namolaru2010a} expresses the space of features via logic programming over relations that represent information from the IRs. It greedily searches for expressions that represent good features. However, their approach relies on expert selected relations, combinators and constraints to work. For both approaches, the search time may be significant.

Cavazos \emph{et al.\ }present a reaction-based predictive model for software-hardware co-design~\cite{Cavazos2006}. Their approach profiles the target program using several carefully selected compiler options to see how program runtime changes under these options for a given micro-architecture setting. They then use the program ``reactions'' to predict the best available application speedup. While their approach does not use static code features, developers must carefully select a few settings from a large number of candidate options for profiling, because poorly chosen options can significantly affect the quality of the model. Moreover, the program must be run several times before optimisation, while our technique does not require the program to be profiled.


% Moren, K., & Gohringer, D. (2018). Automatic Mapping for OpenCL-Programs on CPU/GPU Heterogeneous Platforms. In ICCS. https://doi.org/10.1007/978-3-319-93701-4
\todo[inline]{CPU/GPU mapping~\cite{Shi2018}.}


Deep learning is a nascent branch of machine learning in which deep or multi-level graphs of processing layers are used to detect patterns in natural data~\cite{Buduma2015,LeCun2015}. It is proving especially successful for its ability to ability to process natural data in its raw form. This overcomes the traditionally laborious and time-consuming practise of engineering feature extractors to process raw data into an internal representation or feature vector. Deep learning has successfully discovered structures in high-dimensional data, and is responsible for many breakthrough achievements in machine learning, including human parity in conversational speech recognition~\cite{Xiong2016}; professional level performance in video games~\cite{Mnih2015}; and autonomous vehicle control~\cite{Lozano-Perez2012}.

In past work I used the Long Short-Term Memory (LSTM) architecture of Recurrent Neural Network (RNN)~\cite{Sundermeyer2012,Mikolov2015} to generate sequences of OpenCL code~\cite{Cummins2017a}. The LSTM network architecture comprises recurrent layers of \emph{memory cells}, each consisting of an input, output, and forget gate, and an output layer providing normalised probability values from a 1-of-K coded vocabulary~\cite{Graves,Graves2013}. Although this is the first application of deep learning for generating executable programs, RNNs have been successfully applied to a variety of other generative tasks, including image captioning~\cite{Vinyals}, colourising black and white photographs~\cite{Zhang2016}, artistic style~\cite{Gatys2015}, and image generation~\cite{Gregor2014}.

The proficiency of LSTMs for sequence modelling is demonstrated in~\cite{Sutskever2014}. \citeauthor{Sutskever2014} apply two LSTM networks to translate first a sequence into a fixed length vector, then to decode the vector into an output sequence. This architecture achieves state-of-the-art performance in machine translation. The authors find that reversing the order of the input sequences markedly improves translation performance by introducing new short term dependencies between input and output sequences. Such sequence transformations should be considered for the purpose of program generation.

The application of language modelling for generating executable programs is novel. In training on large corpuses of hand-written code, the language model learns the human biases which are present in common codes~\cite{Caliskan-islam2016}. While such human-induced biases can prove controversial in social domains~\cite{Bolukbasi2016,Joseph2017}, this enables the generation of programs which, unlike other approaches to program generation, are representative of real workloads.

Neural networks are computationally expensive, though their implementations can be generic and parallelised. Library implementations are available in Torch~\cite{Collobert2011}, Caffe~\cite{Jia2014}, and TensorFlow~\cite{Abadi}. The increasing size and depth of computation graphs in deep learning has challenged the ability to compute results in reasonable time. Possible methods for reducing computational overhead involve fusing operations across layers in the graph using domain specific languages~\cite{Truong2016,Ashari2015a,Potter2015}; decoupling interfaces between layers using small networks to synthesise learning gradients during training~\cite{Jaderberg2016a}; and specialising hardware for computing data parallel workloads using FPGAs and ASICs~\cite{Misra2010}.


% \subsubsection{Software engineering}
Machine learning has been applied to source code to aid software engineering.  Naturalize employs techniques developed in the natural language processing domain to model coding conventions~\cite{Allamanis2014a}. JSNice leverages probabilistic graphical models to predict program properties such as identifier names for JavaScript~\cite{Raychev}. \citeauthor{Zhang2015a} use deep learning to generate example code for APIs as responses to natural language queries~\cite{Zhang2015a}. \citeauthor{Allamanis2016} use attentional neural networks to generate summaries of source code~\cite{Allamanis2016}. \citeauthor{Wong2013} mines Q\&A site StackOverflow to automatically generate code comments~\cite{Wong2013}. \citeauthor{Raychev2014} use statistical models to provide contextual code completion~\cite{Raychev2014}.

There is an increasing interest in mining source code repositories at large scale~\cite{Allamanis2013a,White2015a,Bird2009}. Previous studies have involved data mining of GitHub to analyse software engineering practices~\cite{Wu2014,Guzman2014,Baishakhi2014a,Vasilescu2015}; however, no work so far has exploited mined source code for program generation.


% \subsection{Iterative compilation}

Iterative compilation is the method of performance tuning in which a program is compiled and profiled using multiple different configurations of optimisations in order to find the configuration which maximises performance. One of the the first formalised publications of the technique appeared in \citeyear{Bodin1998} by \citeauthor{Bodin1998}~\cite{din1998}.  Iterative compilation has since been demonstrated to be a highly effective form of empirical performance tuning for selecting compiler optimisations.

An enumeration of the optimisation space of Intel Thread Building Blocks in~\cite{Contreras2008} shows that runtime knowledge of the available parallel hardware can have a significant impact on program performance. \citeauthor{Collins2012} exploit this in~\cite{Collins2012}, first using Principle Components Analysis to reduce the dimensionality of the optimisation space, followed by a search of parameter values to improve program performance by a factor of $1.6\times$ over values chosen by a human expert. In~\cite{Collins2013}, they extend this using static feature extraction and nearest neighbour classification to further prune the search space, achieving an average 89\% of the oracle performance after evaluating 45 parameters.

Frameworks for iterative compilation offer mechanisms for abstracting the iterative compilation process from the optimisation space. \emph{OpenTuner} presents a generic framework for optimisation space exploration~\cite{Ansel2013}. \emph{CLTune} is a generic auto-tuner for OpenCL kernels~\cite{Nugteren2015}. Both frameworks implement \emph{search}, however, the huge number of possible compiler optimisations makes such a search expensive to perform for every new configuration of program, architecture and dataset.

% \subsubsection{Machine learning}
Machine learning has been used to guide iterative compilation and predict optimisations for code. \citeauthor{Stephenson2003} use ``meta optimisation'' to tune compiler heuristics through an evolutionary algorithm to automate the search of the optimisation space~\cite{Stephenson2003}. \citeauthor{Fursin2011} continued this with Milepost GCC, the first machine learning-enabled self-tuning compiler~\cite{Fursin2011}. A survey of machine learning heuristics quality concludes that the automatic \emph{generation} of self-tuning heuristics is an ongoing research challenge that offers the greatest generalisation benefits~\cite{Burke2013}.

\citeauthor{Dastgeer2011} developed a machine learning based auto-tuner for the SkePU skeleton library in~\cite{Dastgeer2011}. Training data is used to predict the optimal execution device (i.e.\ CPU, GPU) for a given program by predicting execution time and memory copy overhead based on problem size. The auto-tuner only supports vector operations, and there is limited cross-architecture evaluation. In~\cite{Dastgeer2015a}, the authors extend SkePU to improve the data consistency and transfer overhead of container types, reporting up to a $33.4\times$ speedup over the previous implementation.

\citeauthor{Ogilvie2015} use active learning to reduce the cost of iterative compilation by searching for points in the optimisation space which are close to decision boundaries~\cite{Ogilvie2015}. This reduces the cost of training compared to a random search. \citeauthor{Wahib2015a} use machine learning to automate the selection of GPU kernel transformations~\cite{Wahib2015a}.

% \subsubsection{Dynamic optimisers}
Iterative compilation typically involves searching the optimisation space offline --- dynamic optimisers perform this optimisation space exploration at runtime, allowing optimisations tailored to dynamic feature values. This is a challenging task, as a random search of an optimisation space may result in many configurations with sub-optimal performance. In a real world system, evaluating many sub-optimal configurations will cause significant slowdowns to a program. A resulting requirement of dynamic optimisers is that convergence time towards optimal parameters is minimised.

Existing dynamic optimisation research has typically taken a low level approach to performing optimisations. Dynamo is a dynamic optimiser which performs binary level transformations of programs using information gathered from runtime profiling and tracing~\cite{Bala2000}. While this provides the ability to respond to dynamic features, it restricts the range of optimisations that can be applied to binary transformations. These low level transformations cannot match the performance gains that higher level parameter tuning produces.

% \subsubsection{Superoptimisers}
In~\cite{Massalin1987}, the smallest possible program which performs a specific function is found through an iterative enumeration of the entire instruction set. Starting with a program of a single instruction, the superoptimiser tests to see if any possible instruction passes a set of conformity tests. If not, the program length is increased by a single instruction and the process repeats. The exponential growth in the size of the search space is far too expensive for all but the smallest of hot paths, typically less than 13 instructions. The optimiser is limited to register to register memory transfers, with no support for pointers, a limitation which is addressed in~\cite{Joshi2002}. Denali is a superoptimiser which uses constraint satisfaction and rewrite rules to generate programs which are \emph{provably} optimal, instead of searching for the optimal configuration through empirical testing. Denali first translates a low level machine code into guarded multi-assignment form, then uses a matching algorithm to build a graph of all of a set of logical axioms which match parts of the graph, before using boolean satisfiability to disprove the conjecture that a program cannot be written in $n$ instructions. If the conjecture cannot be disproved, the size of $n$ is increased and the process repeats.

% \subsubsection{GPUs}
Performant GPGPU programs require careful attention from the developer to properly manage data layout in DRAM, caching, diverging control flow, and thread communication. The performance of programs depends heavily on fully utilising zero-overhead thread scheduling, memory bandwidth, and thread grouping. \citeauthor{Ryoo2008a} illustrate the importance of these factors by demonstrating speedups of up to $432\times$ for matrix multiplication in CUDA by appropriate use of tiling and loop unrolling~\cite{Ryoo2008a}. The importance of proper exploitation of local shared memory and synchronisation costs is explored in~\cite{Lee2010}.

In~\cite{Chen2014}, data locality optimisations are automated using a description of the hardware and a memory-placement-agnostic compiler. The authors demonstrate speedups of up to $2.08\times$, although at the cost of requiring accurate memory hierarchy descriptor files for all targeted hardware. The descriptor files must be hand generated, requiring expert knowledge of the underlying hardware in order to properly exploit memory locality.

Data locality for nested parallel patterns is explored in~\cite{Lee}. The authors use an automatic mapping strategy for nested parallel skeletons on GPUs, which uses a custom intermediate representation and a CUDA code generator, achieving $1.24\times$ speedup over hand optimised code on 7 of 8 Rodinia benchmarks.

Reduction of the GPGPU optimisation space is demonstrated in~\cite{Ryoo2008}, using the common subset of optimal configurations across a set of training examples. This technique reduces the search space by 98\%, although it does not guarantee that for a new program, the reduced search space will include the optimal configuration.

\citeauthor{Magni2014} demonstrated that thread coarsening of OpenCL kernels can lead to speedups in program performance between $1.11\times$ and $1.33\times$ in~\cite{Magni2014}. The authors achieve this using a machine learning model to predict optimal thread coarsening factors based on the static features of kernels, and an LLVM function pass to perform the required code transformations.

A framework for the automatic generation of OpenCL kernels from high-level programming concepts is described in~\cite{Steuwer2015}. A set of rewrite rules is used to transform high-level expressions to OpenCL code, creating a space of possible implementations. This approach is ideologically similar to that of PetaBricks, in that optimisations are made through algorithmic choice, although in this case the transformations are performed automatically at the compiler level. The authors report performance on a par with that of hand written OpenCL kernels.

% \subsubsection{Stencils}
Stencil codes have a variety of computationally expensive uses from fluid dynamics to quantum mechanics. Efficient, tuned stencil kernels are highly sought after, with early work in \citeyear{Bolz2003} by \citeauthor{Bolz2003} demonstrating the capability of GPUs for massively parallel stencil operations~\cite{Bolz2003}. In the resulting years, stencil codes have received much attention from the performance tuning research community.

\citeauthor{Ganapathi2009} demonstrated early attempts at autotuning multi-core stencil codes in~\cite{Ganapathi2009}, drawing upon the successes of statistical machine learning techniques in the compiler community. They present an auto-tuner which can achieve performance up to 18\% better than that of a human expert. From a space of 10 million configurations, they evaluate the performance of a randomly selected 1500 combinations, using Kernel Canonical Correlation Analysis to build correlations between tunable parameter values and measured performance targets. Performance targets are L1 cache misses, TLB misses, cycles per thread, and power consumption. The use of KCAA restricts the scalability of their system as the complexity of model building grows exponentially with the number of features. In their evaluation, the system requires two hours of compute time to build the KCAA model for only 400 seconds of benchmark data. They present a compelling argument for the use of energy efficiency as an optimisation target in addition to runtime, citing that it was the power wall that lead to the multi-core revolution in the first place. Their choice of only 2 benchmarks and 2 platforms makes the evaluation of their auto-tuner somewhat limited.

\citeauthor{Berkeley2009} targeted 3D stencils code performance in~\cite{Berkeley2009}. Stencils are decomposed into core blocks, sufficiently small to avoid last level cache capacity misses. These are then further decomposed to thread blocks, designed to exploit common locality threads may have within a shared cache or local memory. Thread blocks are divided into register blocks to take advantage of data level parallelism provided by the available registers. Data allocation is optimised on NUMA systems. The performance evaluation considers speedups of various optimisations with and without consideration for host/device transfer overhead.

\citeauthor{Kamil2010} present an autotuning framework in~\cite{Kamil2010} which accepts as input a Fortran 95 stencil expression and generates tuned shared-memory parallel implementations in Fortan, C, or CUDA. The system uses an IR to explore autotuning transformations, enumerating a subset of the optimisation space and recording only a single execution time for each configuration, reporting the fastest. They demonstrate their system on 4 architectures using 3 benchmarks, with speedups of up to $22\times$ compared to serial implementations. The CUDA code generator does not optimise for the GPU memory hierarchy, using only global memory. As demonstrated in this thesis, improper utilisation of local memory can hinder program performance by two orders of magnitude. There is no directed search or cross-program learning.

In~\cite{Zhang2013a}, \citeauthor{Zhang2013a} present a code generator and auto-tuner for 3D Jacobi stencil codes. Using a DSL to express kernel functions, the code generator performs substitution from one of two CUDA templates to create programs for execution on GPUs. GPU programs are parameterised and tuned for block size, block dimensions, and whether input data is stored in read only texture memory. This creates an optimisation space of up to 200 configurations. In an evaluation of 4 benchmarks, the authors report performance that is comparable with previous implementations of iterative Jacobi stencils on GPUs~\cite{Holewinski2012, Phillips2010}. The dominating parameter is shown to be block dimensions, followed by block size, then read only memory. The DSL presented in the paper is limited to expressing only Jacobi Stencils applications. Their auto-tuner requires a full enumeration of the parameter space for each program, which may be impractical for the needs of general purpose stencil computing. Previous work overcomes this drawback by learning parameter values which transfer to unseen stencils, without the need for an expensive tuning phase for each program and architecture~\cite{Cummins2015}.

In~\cite{Christen2011}, \citeauthor{Christen2011} presents a DSL for expressing stencil codes, a C code generator, and an auto-tuner for exploring the optimisation space of blocking and vectorisation strategies. The DSL supports stencil operations on arbitrarily high-dimensional grids. The auto-tuner performs either an exhaustive, multi-run Powell search, Nelder Mead, or evolutionary search to find optimal parameter values. They evaluate their system on two CPUS and one GPU using 6 benchmarks. The authors do not present a ratio of the available performance that their system achieves, or how the performance of optimisations vary across benchmarks or devices.

A stencil grid can be decomposed into smaller subsections so that multiple GPUs can operate on each subsection independently. This requires a small overlapping region where each subsection meets --- the halo region --- to be shared between devices. For iterative stencils, values in the halo region must be synchronised between devices after each iteration, leading to costly communication overheads. One possible optimisation is to deliberately increase the size of the halo region, allowing each device to compute updated values for the halo region, instead of requiring a synchronisation of shared state. This reduces the communication costs between GPUs, at the expense of introducing redundant computation. Tuning the size of this halo region is the goal of PARTANS~\cite{Lutz2013}, an autotuning framework for multi-GPU stencil computations. \citeauthor{Lutz2013} explore the effect of varying the size of the halo regions using six benchmark applications, finding that the optimal halo size depends on the size of the grid, the number of partitions, and the connection mechanism (i.e.\ PCI express). The authors present an auto-tuner which determines problem decomposition and swapping strategy offline, and performs an online search for the optimal halo size. The selection of overlapping halo region size compliments the selection of work-group size which is the subject of previous work~\cite{Cummins2015}.

In applications of machine learning for iterative compilation, a limiting factor of the effectiveness of learned models is the number of benchmarks used. The development of automatic program generation alleviates this problem by allowing an unbounded number of programs to enumerate the feature space at an increasingly granular scale.


% {Learning Representations}

Recently, deep neural networks~\cite{LeCun2015} have been shown to be a powerful tool for feature engineering in various tasks including image recognition~\cite{Krizhevsky2012,He2016} and audio processing~\cite{Lee2009b}. In the field of compiler optimisation, no work so far has applied deep neural networks for program feature generation and selection. Our work is the first to do so.

%% Wang, H., Raj, B., & Xing, E. P. (2017). On the Origin of Deep Learning, 1–81. Retrieved from http://arxiv.org/abs/1702.07800
Deep Learning is an emerging branch of machine learning, offering breakthrough domains in a number of domains~\cite{Wang2017}.

Automatic inference from raw inputs in other fields. Review of representation learning~\cite{Bengio2013}. Object detection, image classification, text classification~\cite{Conneau2016}, and even to the optimisations themselves~\cite{Andrychowicz2016a}. We are the first to apply this approach to compilers.

End to end training of robotic control policies~\cite{Levine2016}.



% \subsection{Transfer Learning}

%% Razavian, A. S., Azizpour, H., Sullivan, J., & Carlsson, S. (2014). CNN Features off-the-shelf: an Astounding Baseline for Recognition. In CVPRW. https://doi.org/10.1109/CVPRW.2014.131
Exploring relevance of inferred features in model hidden layers~\cite{Yosinski2014}. Transfer learning is common in other domains, e.g. vision~\cite{Razavian2014,Oquab2014}. Often using the hidden layers as fixed feature extractors (by using the output of one sub-model as input to a new model), rather than using the same model structure and transferring weights (as we did).

% Terence, P., & Vinju, J. (2016). Towards a Universal Code Formatter through Machine Learning. In SLE.
\todo[inline]{ML code formatter~\cite{Terence2016}.}

% Allamanis, M. (2018). The Adverse Effects of Code Duplication in Machine Learning Models of Code. ArXiv:1812.06469. Retrieved from https://github.com/Microsoft/dpu-utils.
\todo[inline]{Code duplication~\cite{Allamanis}.}

% Chen, Z., Kommrusch, S., Tufano, M., Pouchet, L., Poshyvanyk, D., & Monperrus, M. (2018). SequenceR: Sequence-to-Sequence Learning for End-to-End Program Repair. ArXiv:1901.01808. Retrieved from http://arxiv.org/abs/1901.01808
\todo[inline]{Program repair~\cite{Chen2018}.}

% Hata, H., Shihab, E., & Neubig, G. (n.d.). Learning to Generate Corrective Patches using Neural Machine Translation. ArXiv:1812.07170.
\todo[inline]{Corrective patches~\cite{Hata}.}

% Ben-nun, T., Jakobovits, A. S., & Hoefler, T. (2018). Neural Code Comprehension: A Learnable Representation of Code Semantics. In NeurIPS.
\todo[inline]{Neural Code Comprehension~\cite{Ben-nun2018} builds on the experiments in Chapter~\ref{chap:deeptune} of this thesis, using a novel XFG representation. Assembled from LLVM byte-code so language portable.}

% Yin, P., Neubig, G., Allamanis, M., Brockschmidt, M., & Gaunt, A. L. (2018). Learning to Represent Edits. ArXiv:1810.13337. https://doi.org/10.1080/02688697.2016.1244252
\todo[inline]{Learning to represent edits and diffs~\cite{Yin2018}~\cite{Tufano2019}.}

% Wang, K., Singh, R., & Su, Z. (2017). Dynamic Neural Program Embeddings for Program Repair. ArXiv:1711.07163.
\todo[inline]{Program embeddings~\cite{Wang2017d}.}

% Henkel, J., Lahiri, S. K., Liblit, B., & Reps, T. (2018). Code Vectors: Understanding Programs Through Embedded Abstracted Symbolic Traces. In FSE. https://doi.org/10.1145/3236024.3236085
\todo[inline]{Embeddings from AST~\cite{Henkel2018}.}


\todo[inline]{Integrate PACT'17 Related work section}


Machine learning has emerged as a viable means in automatically constructing heuristics for code optimisation~\cite{Wang2010,Kulkarni2012,Muralidharan2016,Ogilvie2017,Ren,Cummins2016}. Its great advantage is that it can adapt to changing hardware platforms as it has no a priori assumptions about their behaviour. The success of machine learning based code optimisation has required having a set of high-quality features that can capture the important characteristics of the target program. Given that there is an infinite number of these potential features, finding the right set of features is a non-trivial, time-consuming task.

Various forms of program features have been used in compiler-based machine learning. These include static code structures~\cite{Jiang2010} and runtime information such as system load~\cite{Wen2015} and performance counters~\cite{Dubach2009}. In compiler research, the feature sets used for predictive models are often provided without explanation and rarely is the quality of those features evaluated. More commonly, an initial large, high dimensional candidate feature space is pruned via feature selection~\cite{Stephenson2005,Taylor}, or projected into a lower dimensional space~\cite{Collins2013,Dubach2007}. FEAST employs a range of existing feature selection methods to select useful candidate features~\cite{Ting2016}. Unlike these approaches, DeepTune extracts features and reduces the dimensionality of the feature space completely internally and without expert guidance.

Park \emph{et al.} present a unique graph-based approach for feature representations~\cite{Park2012}. They use a Support Vector Machine where the kernel is based on a graph similarity metric. Their technique still requires hand coded features at the basic block level, but thereafter, graph similarity against each of the training programs takes the place of global features. Being a kernel method, it requires that training data graphs be shipped with the compiler, which may not scale as the size of the training data grows with the number of instances, and some training programs may be very large. Finally, their graph matching metric is expensive, requiring $O(n^3)$ to compare against each training example. By contrast, our method does not need any hand built static code features, and the deployment memory footprint is constant and prediction time is linear in the length of the program, regardless of the size of the training set.

A few methods have been proposed to automatically generate features from the compiler's intermediate representation~\cite{Namolaru2010a,Leather2014}. These approaches closely tie the implementation of the predictive model to the compiler IR, which means changes to the IR will require modifications to the model. The work of \cite{Leather2014} uses genetic programming to search for features, and required a huge grammar to be written, some 160kB in length. Although much of this can be created from templates, selecting the right range of capabilities and search space bias is non trivial and up to the expert. The work of \cite{Namolaru2010a} expresses the space of features via logic programming over relations that represent information from the IRs. It greedily searches for expressions that represent good features. However, their approach relies on expert selected relations, combinators and constraints to work. For both approaches, the search time may be significant.

Cavazos \emph{et al.\ }present a reaction-based predictive model for software-hardware co-design~\cite{Cavazos2006}. Their approach profiles the target program using several carefully selected compiler options to see how program runtime changes under these options for a given micro-architecture setting. They then use the program ``reactions'' to predict the best available application speedup. While their approach does not use static code features, developers must carefully select a few settings from a large number of candidate options for profiling, because poorly chosen options can significantly affect the quality of the model. Moreover, the program must be run several times before optimisation,while our technique does not require the program to be profiled.

In recent years, machine learning techniques have been employed to model and learn from program source code on various tasks. These include mining coding conventions~\cite{Allamanis2014a} and idioms~\cite{Allamanis2014}, API example code~\cite{Zhang2015a} and pseudo-code generation~\cite{Oda2015}, and benchmark generation~\cite{Cummins2017a}. Our work is the first attempt to extend the already challenging task of modelling distributions over source code to learning distributions over source code with respect to code optimisations.

Recently, deep neural networks have been shown to be a powerful tool for feature engineering in various tasks including image recognition~\cite{Krizhevsky2012,He2016} and audio processing~\cite{Lee2009b}. No work so far has applied deep neural networks for program feature generation. Our work is the first to do so.

\section{Other Machine Learning}
\label{sec:related-work-other}

There is an increasing interest in mining source code repositories at large scale~\cite{Allamanis2013a,White2015a,Bird2009}. Previous studies have involved data mining of GitHub to analyse software engineering practices~\cite{Wu2014,Guzman2014,Baishakhi2014a,Vasilescu2015}, for example code generation~\cite{Zhang2015a}, code summarisation~\cite{Allamanis2016}, comment generation~\cite{Wong2013}, and code completion~\cite{Raychev2014}. However, no work so far has exploited mined source code for benchmark generation. This work is the first to do so.

In recent years, machine learning techniques have been employed to model and learn from program source code on various tasks. These include mining coding conventions~\cite{Allamanis2014a} and idioms~\cite{Allamanis2014}, API example code~\cite{Zhang2015a} and pseudo-code generation~\cite{Oda2015}. Our work is the first attempt to extend the already challenging task of modelling distributions over source code to learning distributions over source code with respect to code optimisations.

In a recent work~\cite{Bunel2017a}, Bunel \emph{et al.} formulate code super-optimisation as a stochastic search problem to learn the distribution of different code transformations and expected performance improvement. As acknowledged by the authors, their approach can be improved by having temporal information of the code structures.

% ML FOR TESTING:

Machine learning has also been applied to other areas such as improving bug
finding static analysers~\cite{Heo2017,Koc2017}, repairing
programs~\cite{Koukoutos2017a,White}, prioritising test
programs~\cite{Chen2017}, identifying buffer overruns~\cite{Choi2016}, and
processing bug reports~\cite{Lam2016,Huo2016}. To the best of our knowledge, no
work so far has succeeded in finding compiler bugs by exploiting the learned
syntax of mined source code for test case generation. Ours is the first to do
so.