% Education.

\Title{Education}

\Description{%
  % Date graduated.
  \MarginText{2019\\(expected)}%
  % Degree title.
  \textbf{Ph.D, Informatics}
  \newline
  % School.
  University of Edinburgh, School of Informatics
  \vspace{.5em}\\ \noindent
  % Description.
  Deep learning over programs. Developing novel machine learning methods for random program generation, compiler optimisations, and representative benchmarking. Applications for heterogeneous parallelism, compiler testing, and adaptive performance tuning. To date: 6 publications, 2 best papers, 7 invited talks, 5 posters, 4 conferences.%
  % Conferences: HiPEAC'16, PLDI'16, CGO'17, PACT'17.
}


\Description{%
  % Date graduated.
  \MarginText{2015}%
  % Degree title.
  \textbf{MSc by Research, Pervasive Parallelism \textit{(Distinction)}}
  \newline
  % School.
  University of Edinburgh, School of Informatics
  \newline
  % Thesis.
  Thesis: \textit{Autotuning Stencil Codes with Algorithmic Skeletons} (grade: 85\%)
  \vspace{.5em}\\ \noindent
  % Description.
  Runtime adaptive tuning for heterogeneous parallel systems, targeting a high level DSL for multi-GPU stencil programs. Machine learning over distributed training sets.%
}


\Description{%
  % Date graduated.
  \MarginText{2014}%
  % Degree title.
  \textbf{MEng Electronic Engineering \& Computer Science \textit{(First Class Honours)}}
  \newline
  % School.
  Aston University, School of Engineering \& Applied Science
  \newline
  % Thesis.
  Thesis: \textit{Protein Isoelectric Point Database} (grade: 90\%)
  \vspace{.5em}\\ \noindent
  % Description.
  Created a search engine and API for a novel molecular biochemistry dataset. Targeting bioinformatics research and released open source, with peak 854 monthly active users.%  Source: Google Analytics July 1-July 31 2015.
}
