% Jobs.

\Title{Professional Experience}

\Description{%
% Date range.
\MarginText{2018\\*\textit{(current)}}%
% Company.
\textbf{DeepMind}, London, UK
\newline%
% Job title.
Software Engineer Intern
\vspace{.5em}\\ \noindent
% Description.
Investigating deep reinforcement learning for optimizing instruction fusion of TensorFlow XLA graphs. Black-box learning methods for TPU, GPU, and CPU performance.%
}

\Description{%
% Date range.
\MarginText{2018}%
% Company.
\textbf{Google}, Mountain View, California, USA
\newline%
% Job title.
Software Engineer Intern
\vspace{.5em}\\ \noindent
% Description.
Created a representative benchmark of Google's Protocol Buffer usage. Working in the Google Wide Profiling team to synthesise benchmarks of Google compute. The project involved company-wide workload characterization through to datacentre-scale low level performance analysis of profiles and hardware counters.%
}


\Description{%
% Date range.
\MarginText{2016--2018}%
% Company.
\textbf{Codeplay Software}, Edinburgh, UK
\newline%
% Job title.
Software Engineer, Part Time
\vspace{.5em}\\ \noindent
% Description.
Developing OpenCL port of Tensorflow. Implemented GPU memory management for Eigen. Compile time scheduling and kernel fusion for expression trees on GPUs. Proposed and designed a Python interface for VisionCpp as lead developer. Extensive C++ meta-programming. Integrated compiler fuzzing into continuous testing tooling.%
}


\Description{%
% Date range.
\MarginText{2012--2013}%
% Company.
\textbf{Intel Corporation}, London, UK
\newline%
% Job title.
Open Source Developer Intern
\vspace{.5em}\\ \noindent
% Description.
Patched \texttt{ioctl} subsystem in Linux kernel. Developed a novel SIMD register visualisation tool for Intel GPU assembly programming. Implemented GTK+ support for Wayland display server. Fixed memory and usability bugs in GNOME desktop applications. Developed particle effects engine for a 3D rendering program. Rapid prototyping of Android applications. Numerous contributions to open source projects.%
}

%\newpage
%\Description{%
%  % Date range.
%  \MarginText{2010--2017}%
%  % Company.
%  \textbf{Freelance Web Developer}
%	\newline
%  % Description.
%  Full-stack development for small businesses, including graphic design and branding. Clients included conferences, publishing houses, musicians, and a beauty parlour.%
%  % Clients:
%  %
%  % CGO'18              http://cgo.org/cgo2018/
%  % The Beauty Rooms    http://beautyroomsmatlock.com
%  % Barthelemy Jusselme http://barthelemyjusselme.com
%  % Myrmidon Books      http://myrmidonbooks.com
%  % Border Scripts      http://borderscripts.com
%  % Greystones Press    http://www.greystonespress.com
%}


%\Description{%
%  % Date range.
%  \MarginText{2008}%
%  % Company.
%  \textbf{Rolls Royce Holdings plc}, Derby, UK
%	\newline
%  % Description.
%  Work placement in the Design Methods \& Improvements team.%
%}
